---
title: "Case 2 Part 3"
author: "Sonia Xu, Grant Goettel, Ian Hua"
date: "October 27, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r echo = F, warning = F, include = F}
knitr::opts_chunk$set(fig.width=5, fig.height=3) 
library(dplyr)
library(magrittr)
library(knitr)
library(ggplot2)
```


```{r include = F}
kelly <- read.table("kellydat.txt", header = T)
kelly <- kelly %>% mutate(count_sn = sn1 + sn2*2 + sn3*3 + all4*4) %>% select(-c(sn1, sn2, sn3, all4))
```

<!-- BACKGROUND: Study of time to critical neurological assessment for patients with stroke-like symptoms who are admitted to the emergency room.  A possible predictor is number of major stroke symptoms reported, ranging from 0 to 4.  Treatment for acute stroke includes thrombolytic therapy, which can potentially improve neurological functioning for ischemic stroke patients if administered soon after symptom onset (within 3 hours).  Since treating patients quickly is critically important for their long-term prognosis, minimizing the times from symptom onset to emergency room (ED) arrival, from ED arrival to diagnosis, and from diagnosis to treatment is of paramount concern.
INTEREST: Factors predictive of the time to critical neurological assessment following admission to the ED for n=335 patients with mild to moderate motor impairment.  The goal of the analysis is to perform inferences on the impact of clinical presentation, gender and race on time to neurological assessment.  Clinical presentation is measured as a count of reported major stroke symptoms, including headache, loss of motor skills or weakness, trouble talking or understanding, and vision problems.
-->
#Introduction
The following data is from a study of time to critical neurological assessment for patients with stroke-like symptoms who were admitted to the emergency room. The purpose of the analysis is to perform inferences on the impact of clinical presentation (reported number of makor stroke symptoms), gender, and race on time to neurological asessment. This paper fits a model that identifies the differences in wait time to neurological assessment based on these features of interest. A final model of kernel regression with a response of time at failure, continuous feature of nctdel estimated via kernel regression, and categorical features of black, hispanic, male, number of symptoms.

#Dataset
The dataset contained information about the amount of time elapsed from arrival at the ER to the assessment, and whether or not the patient received a CT scan. It also contained data about whether each patient was male or female, whether they were black or hispanic or not, and the amount of major symptoms of a stroke they had upon arrival (out of 4 main symptoms). Rather than have four separate indicator variables showing whether they had 1, 2, 3, or 4 symptoms, we created a single numerical variable containing the amount of symptoms.

#Methodology
To assess the impact of clinical presentation, gender and race on time to neurological assessment, multiple models and analyses were explored to find the best model that fit the data. After exploring Cox Proportional Hazards, Kaplan-Meier Estimate, Random Forest, and Kernel Regression, the Kernel Regression Model seemed to perform the best despite its flaws. We chose 10 bins based on quantiles to account for the uneven frequencies of the data. 

#EDA

##Racial Bias
###Black vs. Non-Black
There could potentially be bias here, but more analysis should be done. Non-blacks appear to have shorter wait times, but some non-blacks have extremely long wait times.

```{r echo = F}

ggplot(kelly, aes(nctdel, group = factor(black), fill = factor(black,labels=c("no", "yes"))),alpha=0.5, adjust=2) + geom_density() + theme_bw() + xlab("Wait Time") + ylab("Frequency of Wait Time") + labs(fill = "Black?") + ggtitle("Differences in Wait Time if Black")
```

###Hispanic vs. Non-Hispanic
There could potentially be bias here, but the lack of sample size for Hispanics appears to be an issue.
```{r echo = F}
ggplot(kelly, aes(nctdel, group = factor(hisp), fill = factor(hisp,labels=c("no", "yes"))),alpha=0.5, adjust=2) + geom_density() + theme_bw() + xlab("Wait Time") + ylab("Frequency of Wait Time") + labs(fill = "Hispanic?") + ggtitle("Differences in Wait Time if Hispanic")
```


##Gender Bias
There could potentially be bias here, but more analysis should be done. Females appear to have shorter wait times.

```{r echo = F}
ggplot(kelly, aes(nctdel, group = factor(male), fill = factor(male,labels=c("no", "yes"))),alpha=0.5, adjust=2) + geom_density() + theme_bw() + xlab("Wait Time") + ylab("Frequency of Wait Time") + labs(fill = "Male?") + ggtitle("Differences in Wait Time if Male")
```

#Final Model

We decided to create a Kernel Regression model with time at failure as the response to explore the robustness of this new model.

##Kernel Regression with 10 Bins
10 bins were calculated to fit the kernels. The bins are unevenly spaced because the data has a higher concentration of points for the feature nctdel between 0 and 2, even though its range is (0,26.25). We chose 10 bins because we wanted to be careful of the sample size with only 335 observations. If we split that over too many bins, the sample sizes in the individual kernels would be uncomfortably small for regression. The model has time at failure as the response, a kernel estimation of nctdel, and the features male, black, hispanic, and count_sn. The bin levels are (-Inf,0], (0,0.357], (0.357,0.737], (0.737,0.983], (0.983,1.21], (1.21,1.49], (1.49,1.73], (1.73,2.16], (2.16,2.8], (2.8, Inf]. A summary of the model noted some significance for the feature nctdel (for the full summary, Appendix A).

##Bins

```{r echo = F}
# set up the kernels
#bin <- cut(kelly$nctdel, c(-Inf,seq(0,26,.3),Inf))
quants <- quantile(kelly$nctdel, probs = seq(0,1,length.out = 10))
bin <- cut(kelly$nctdel, c(-1,quants))
levels(bin) <- seq(1:nlevels(bin))
```


```{r, echo = F, warning = F}
# instantiate data
times <- as.numeric(bin)
total_bins <- max(times)
censored <- kelly$fail
charac <- kelly %>% select(-c(nctdel, fail))
# example transformation
# transformation function

transformation <- function(time, censored, total_bins) {
  y <- rep(0,time)
  if(!censored) {
    y[time] <- 1
  }
  
  # if you are doing kernel regression then you will calculate X differently here (using the kernels)
  X <- array(0, dim=c(time, total_bins))
  diag(X) <- 1
  
  return(data.frame(y,X))
}
# apply the transformation to the data set
d <- transformation(times[1], censored[1], total_bins)
for(i in 2:length(times)) {
  d <- rbind(d, transformation(times[i], censored[i], total_bins))
}
# fit a model
#m <- glm(y ~ 0 + ., data=d, family="binomial")
#summary(m)

#beta <- coef(m)
#hazard <- exp(beta)/(1+exp(beta))
#plot(1:total_bins, hazard, type='l')

# this is pretty ugly. Let's smooth it out with kernel regression.
# set up the kernels
kernels <- 4
tau <- seq(from=1, to=total_bins, length.out = kernels)
sigma <- (tau[2]-tau[1])/2
# we're going to pre-calculate a bunch of kernel weights
# each row is one bin, the columns are the kernel weights for that bin
kernel.weights <- matrix(dnorm(rep(1:10,kernels),rep(tau,each=10),sigma), ncol=kernels)
# now we'll create a new transform function
kernel.transformation <- function(time, censored, other_cov, total_bins) {
  y <- rep(0,time)
  if(!censored) {
    y[time] <- 1
  }
  
  # X has kernel weights instead of the bin indicators that it had before
  X <- array(kernel.weights[1:time,],dim=c(time,kernels))
  X <- cbind(X, other_cov)
  
  return(data.frame(y,X))
}

d2 <- NULL
d2 <- kernel.transformation(times[1], censored[1], charac[1,],total_bins)
for(i in 2:length(times)) {
  d2 <- rbind(d2, kernel.transformation(times[i], censored[i], charac[i,], total_bins))
}

# fit a model
m2 <- glm(y ~ 0 + ., data=d2, family="binomial")
```


##Model Checks
To understand how well the model fits, we performed goodness of fit tests and a model check assumptions.
```{r echo = F}
par(mar=c(1.5,1.5,1.5,1.5))
par(mfrow = c(2,2))
plot(m2)
```



Looking at the Residuals vs. Fitted graph, the points do not exhibit a pattern and are relatively evenly distributed. With the Residuals vs. Leverage plot, the points are not homoscedastically distributed. In addition, data no longer tends to be normal the longer the patient stays waiting, as shown in the Normal Q-Q plot. Generally, the model seems to decently fit our data and is exceptionally better than previous models.
```{r echo = F}
X.test <- data.frame(d2) %>% select(-y)
kern.pred <- round(predict.glm(m2, newdata = data.frame(d2), type = 'response'))
m = mean(kern.pred == d2$y)
```
The model fits the true dataset `r m*100`% of the time when predicting for failure over the entire dataset, implying a decent model which holds much potential given the assumptions hold.


##Discussion

##Gender
Looking at the survival function of Males (with no symptoms, neither black nor male) against Females (with the same characteristics), it reaffirms the EDA that females tend to have a lower time until assessment than males. 

```{r echo = F}
knitr::opts_chunk$set(fig.width=5, fig.height=3) 
##select rows in d2
##first one to test is only black--> use as baseline?

haz_charac_base <- charac[1:10,] %>% mutate(male = 0, black = 0, hisp = 0, count_sn = 0) #female, not black or hispanic, no symptoms
haz_charac_1 <- charac[1:10,] %>% mutate(male = 1, black = 0, hisp = 0, count_sn = 0)

hazard.logodds_base <- predict(m2, newdata = data.frame(cbind(kernel.weights, haz_charac_base)))
hazard.logodds_1 <- predict(m2, newdata = data.frame(cbind(kernel.weights, haz_charac_1)))
###change to assess different survival functions

hazard2_base <- exp(hazard.logodds_base)/(1+exp(hazard.logodds_base))
hazard2_1 <- exp(hazard.logodds_1)/(1+exp(hazard.logodds_1))
hazard2_1 <- hazard2_1*(1-hazard2_base)

plot(1:total_bins, hazard2_base, type='l', col = "black", ylim=(c(0,0.06)), ylab = "hazard", xlab = "time bins")
par(new=T)
plot(1:total_bins, hazard2_1, col = "red",type='l', ylim=(c(0,0.06)), xlab = "time bins", ylab = "hazard", main = "Males vs. Females", sub = "With No Symptoms, Neither Black nor Hispanic")
legend(8,0.05, legend=c("Male", "Female"), col=c("black", "red"), lty = c(1,1), cex=0.5)
```


##Race

###Black
Again, this survival function reaffirms that Non-Black patients have shorter wait times than Black patients (with features 4 symptoms, male, and non-hispanic held constant).

```{r echo = F}
##select rows in d2
##first one to test is only black--> use as baseline?

haz_charac_base <- charac[1:10,] %>% mutate(male = 1, black = 1, hisp = 0, count_sn = 4) #male, black,4 symptoms
haz_charac_1 <- charac[1:10,] %>% mutate(male = 1, black = 0, hisp = 0, count_sn = 4)

hazard.logodds_base <- predict(m2, newdata = data.frame(cbind(kernel.weights, haz_charac_base)))
hazard.logodds_1 <- predict(m2, newdata = data.frame(cbind(kernel.weights, haz_charac_1)))
###change to assess different survival functions

hazard2_base <- exp(hazard.logodds_base)/(1+exp(hazard.logodds_base))
hazard2_1 <- exp(hazard.logodds_1)/(1+exp(hazard.logodds_1))
hazard2_1 <- hazard2_1*(1-hazard2_base)

plot(1:total_bins, hazard2_base, type='l', col = "black", ylim=(c(0,0.06)), ylab = "hazard", xlab = "time bins")
par(new=T)
plot(1:total_bins, hazard2_1, col = "red",type='l', ylim=(c(0,0.06)), xlab = "time bins", ylab = "hazard", main = "Black vs Non-Black", sub = "With 4 Symptoms, Male, Non-Hispanic")
legend(8,0.05, legend=c("Black", "Non-Black"), col=c("black", "red"), lty = c(1,1), cex=0.5)
```

###Hispanic

Hispanic patients tend to have a shorter wait time than Non-Hispanic patients.

```{r echo = F}
##select rows in d2
##first one to test is only black--> use as baseline?

haz_charac_base <- charac[1:10,] %>% mutate(male = 1, black = 0, hisp = 1, count_sn = 4) #male, black,4 symptoms
haz_charac_1 <- charac[1:10,] %>% mutate(male = 1, black = 0, hisp = 0, count_sn = 4)

hazard.logodds_base <- predict(m2, newdata = data.frame(cbind(kernel.weights, haz_charac_base)))
hazard.logodds_1 <- predict(m2, newdata = data.frame(cbind(kernel.weights, haz_charac_1)))
###change to assess different survival functions

hazard2_base <- exp(hazard.logodds_base)/(1+exp(hazard.logodds_base))
hazard2_1 <- exp(hazard.logodds_1)/(1+exp(hazard.logodds_1))
hazard2_1 <- hazard2_1*(1-hazard2_base)

plot(1:total_bins, hazard2_base, type='l', col = "black", ylim=(c(0,0.06)), ylab = "hazard", xlab = "time bins")
par(new=T)
plot(1:total_bins, hazard2_1, col = "red",type='l', ylim=(c(0,0.06)), xlab = "time bins", ylab = "hazard", main = "Hispanic vs Non-Hispanic", sub = "With 4 Symptoms, Male, Non-Black")
legend(8,0.05, legend=c("Hispanic", "Non-Hispanic"), col=c("black", "red"), lty = c(1,1), cex=0.5)
```

##Symptoms

Patients who exhibit 4 symptoms have shorter wait times than those who have 3,2, and 1. 

```{r echo = F}
##select rows in d2
##first one to test is only black--> use as baseline?

haz_charac_base <- charac[1:10,] %>% mutate(male = 1, black = 0, hisp = 0, count_sn = 4) #male, black,4 symptoms
haz_charac_1 <- charac[1:10,] %>% mutate(male = 1, black = 0, hisp = 0, count_sn = 3)
haz_charac_2 <- charac[1:10,] %>% mutate(male = 1, black = 0, hisp = 0, count_sn = 2)
haz_charac_3 <- charac[1:10,] %>% mutate(male = 1, black = 0, hisp = 0, count_sn = 1)

hazard.logodds_base <- predict(m2, newdata = data.frame(cbind(kernel.weights, haz_charac_base)))
hazard.logodds_1 <- predict(m2, newdata = data.frame(cbind(kernel.weights, haz_charac_1)))
hazard.logodds_2 <- predict(m2, newdata = data.frame(cbind(kernel.weights, haz_charac_2)))
hazard.logodds_3 <- predict(m2, newdata = data.frame(cbind(kernel.weights, haz_charac_3)))
###change to assess different survival functions

hazard2_base <- exp(hazard.logodds_base)/(1+exp(hazard.logodds_base))
hazard2_1 <- exp(hazard.logodds_1)/(1+exp(hazard.logodds_1))
hazard2_1 <- hazard2_1*(1-hazard2_base)
hazard2_2 <- exp(hazard.logodds_2)/(1+exp(hazard.logodds_2))
hazard2_2 <- hazard2_2*(1-hazard2_base)*(1-hazard2_1)
hazard2_3 <- exp(hazard.logodds_3)/(1+exp(hazard.logodds_3))
hazard2_3 <- hazard2_3*(1-hazard2_base)*(1-hazard2_1)*(1-hazard2_2)

plot(1:total_bins, hazard2_base, type='l', col = "black", ylim=(c(0,0.06)), ylab = "hazard", xlab = "time bins")
par(new=T)
plot(1:total_bins, hazard2_1, type='l', col = "red", ylim=(c(0,0.06)), ylab = "hazard", xlab = "time bins")
par(new=T)
plot(1:total_bins, hazard2_2, type='l', col = "yellow", ylim=(c(0,0.06)), ylab = "hazard", xlab = "time bins")
par(new = T)
plot(1:total_bins, hazard2_3, col = "green",type='l', ylim=(c(0,0.06)), xlab = "time bins", ylab = "hazard", main = "Symptoms", sub = "With Male, Non-Black, Non-Hispanic")

legend(8,0.05, legend=c("4 Symptoms", "3 Symptoms", "2 Symptoms", "1 Symptom"), col=c("black", "red", "yellow", "green"), lty = c(1,1), cex=0.5)
```

#Recommendations
Based on the survival function produced from the fitted model, there appears to be a positive bias towards females, Black patients, and non-Hispanic patients. In a perfect world, assesssment bias should only depend on the number of symptoms--intuitively, the more symptoms, the greater urgency, so a shorter wait time. Further qualitative studies should be conducted to understand why neurological assessments favor females, Black, and non-Hispanic patients.

A kernel regression hazard model is dependent on all the observations; a slight change in some observations with high leverage could potentially change the model quite a bit. Reproducing this study on various differing datasets could prove beneficial in discovering a better model. It would be also be interesting to test different percentiles for the binning of the model. <!---In addition, the bump in the hazard plot seems quite peculiar and the cases for those patients should be investigated further.-->

#Appendix A
```{r echo = F}
summary(m2)
```
