---
title: "Case 2 Part 3"
author: "Sonia Xu"
date: "October 22, 2017"
output: html_document
---

```{r echo = F, warning = F, include = F}
knitr::opts_chunk$set(fig.width=5, fig.height=3) 
library(dplyr)
library(magrittr)
library(knitr)
library(ggplot2)
```


```{r include = F}
kelly <- read.table("kellydat.txt", header = T)
kelly <- kelly %>% mutate(count_sn = sn1 + sn2*2 + sn3*3 + all4*4) %>% select(-c(sn1, sn2, sn3, all4))
```

#Introduction
```{r}

```

#Methodology
```{r}

```

#Dataset
The data set contained information about the amount of time elapsed from arrival at the ER to the assessment, and whether or not the patient received a CT scan. It also contained data about whether each patient was male or female, whether they were black or hispanic or not, and the amount of major symptoms of a stroke they had upon arrival (out of 4 main symptoms). Rather than have four separate indicator variables showing whether they had 1, 2, 3, or 4 symptoms, we created a single numerical variable containing the amount of symptoms.



#Final Model
```{r}
```

##EDA
```{r}
```

##Goodness of Fit Test
```{r}
```

##Discussion
```{r}
```

